---
title: "metcorrection"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Replace missing values from required data columns


First read in - you will need to replace name to your local versions
**note changing 05 in the name to 5 is necessary**

Consider making a back up of the original file as it will be modified here
```{r read}
Kc_2017_data_60 <- read.table("Kc_2017_data_60.txt", header=T)
knew = Kc_2017_data_60
```

Modify
```{r mod}

# windspeed is required to you have to get rid of the -9999
knew = knew %>% mutate(U = ifelse(U < 0, 0, U))

# get rid of zero windspeeds as this causes an error in estimating aerodynamic conductance
knew = knew %>% mutate(U = ifelse(U == 0, 0.1, U))
# pressure is not in correct units
# I think it is inch of mecury should be kPa - multiply by 3.386

knew = knew %>% mutate(pres = pres*3.386)

knew = knew %>% mutate(pres = ifelse(pres == 0, 98, pres))

knew = knew %>% mutate(kdown = ifelse(kdown <0 , 0, kdown))

knew = knew %>% mutate(RH = ifelse(RH == 0, 47, RH))

knew = knew %>% mutate(xsmd = ifelse(xsmd > 0, -999, xsmd))
 
# check all of the required values
summary(knew[,c("U","RH","Tair","pres","rain","kdown")])
```
When happy overwrite file
```{r mod}
write.table(knew, file="Kc_2017_data_60.txt", row.names=F)
```

## Site Select Issues

Something looks weird about zo and zd - (and I think this causes the following error)
So I changed them as well
to 0.01 0.2 respectively


