---
title: "SUEWS"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```
*All of these graphs are comparing air temperature at 2 m above ground level*



This markdown for visualizing initial SUEWS runs
Full pavement vs full trees and grass
```{r}
pavement <- read.csv("pavement.txt", sep="")

all_trees <- read.csv("all_trees.txt", sep= "")

more_tree <- read.csv("moretree.txt", sep= "")

```
Wrangle data so that it only shows daily high temp 
```{r}
daily_high_all_trees <- all_trees %>% 
  group_by(DOY) %>%
  filter(Tsurf == max(Tsurf)) %>% 
  distinct(DOY, .keep_all = TRUE)

daily_high_pavement <- pavement %>% 
   group_by(DOY) %>%
  filter(Tsurf == max(Tsurf)) %>% 
  distinct(DOY, .keep_all = TRUE)
```

Merge dataframes 
```{r}
merged_test <- merge(daily_high_pavement,daily_high_all_trees,by='DOY', all.x=T) %>% 
  select(DOY, Year.x, Hour.x, T2.x, T2.y) %>% 
    rename(
    Year = Year.x, 
    Hour = Hour.x,
    Before = T2.x,
    After = T2.y
    )
  
```

Graph the temps
```{r}
graph <- ggplot(merged_test, aes(x = DOY)) +
  geom_line(aes(y = Before), color = "red") +
  geom_line(aes(y = After), color = "green") +
  scale_color_manual(values = c("#000033", "#00AFBB"))
  

graph


```


Visualizing runs with semi realistic land cover fractions (0.3 paved, 0.2 buildings, 0.1 grass, 0.1 soil, 0.3 dec trees)
Changed from 0 irrigation to full irrigation

```{r}
no_irrigation <- read.csv("no_irrigation.txt", sep="")



irrigation <- read.csv("irrigation.txt", sep = "")

```

```{r}
names(pavement)
summary(pavement$Tsurf)
summary(all_trees$Tsurf)
summary(pavement$Irr)
summary(all_trees$Irr)
summary(pavement$LAI)
summary(all_trees$LAI)
summary(more_tree$T2)
summary(no_irrigation$T2)


plot(all_trees$Evap*24)

```

```{r}
daily_high_no_irr <- no_irrigation %>% 
  group_by(DOY) %>%
  filter(Tsurf == max(Tsurf)) %>% 
  distinct(DOY, .keep_all = TRUE)

daily_high_irr <- irrigation %>% 
   group_by(DOY) %>%
  filter(Tsurf == max(Tsurf)) %>% 
  distinct(DOY, .keep_all = TRUE)

```

```{r}
merged_test2 <- merge(daily_high_no_irr,daily_high_irr,by='DOY', all.x=T) %>% 
  select(DOY, Year.x, Hour.x, T2.x, T2.y) %>% 
    rename(
    Year = Year.x, 
    Hour = Hour.x,
    Before = T2.x,
    After = T2.y
    )
```

```{r}
graph2 <- ggplot(merged_test2, aes(x = DOY)) +
  geom_line(aes(y = Before), color = "darkred") +
  geom_line(aes(y = After), color = "green") 
 

graph2
```

