---
title: "GP_UHI"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

***ISSUES*** 
 The function is not solved for x. With fixed values, I can use uniroot to find the solution, but this doesnt't work when I try to use values from a dataframe. 
 
Line 72 is my latest attempt at using our dataframe
*STEP 0* 

Load packages 
```{r}
library(rootSolve)
library(tidyverse)
library(purrr)
library(dplyr)
library(MASS)
library(reshape)

```


load csv 
```{r}
uhi <- read_csv("uhi_test.csv")
hourly <- read_csv("hourly_2017_met.csv")

```

### Surface Energy Balance - Solving for surface temperature to get a sense of how much surface temp changes with irrigation

*STEP 1*

A. Run this code first to set the variables.

```{r vars}
T_a = 33 # ambient temp in Celsius 
alpha = 0.2 # albedo 
eps = 0.95 # emissivity 
sigma = 5.6704*10^-8 #W/m^2/K^4 stefan-boltzmann constant
k_H = 21 # sensible heat conductivity [J/deg-C]
k_E = 0 # 0 for non-wetted surface, 100 for fully wet. [J/kPa] Latent energy only present if surface is wet

RH = 0.7 # Relative humidity of atmosphere, [0-1]
RH_s = 0 # Relative humidity of the surface (assume 1 is wet/irrigated, 0 is dry)

K_inc = 750 # W/m^2, incoming shortwave radiation
#L_inc = eps * sigma * T_a^4  # W/m^2, incoming longwave radiation
Q_av = 0.9 * (K_inc*(1-alpha) + eps * sigma * T_a^4) # W/m^2 

```

B. Run this to solve for the surface temperature (with fixed values)

```{r function}
# temperature is represented as x 
fun <- function(x){ (-Q_av + eps*sigma*(x+273.15)^4 + k_H*(x-T_a) + k_E*(RH_s*0.61078*exp(17.27*x/(x+237.3)) - RH*0.61078*exp(17.27*T_a/(T_a+237.3))))

}

# solve function for temp, within the set range 
range = c(0,100)
temp_s = uniroot.all(fun, range)
temp_s


```


C. Try inputting dataframe uhi to function 

```{r function}
# temperature is represented as x 
fun <- function(alpha, eps, sigma, RH, RH_s, T_a, K_inc, k_H, k_E, Q_av){ (-Q_av + eps*sigma*(x+273.15)^4 + k_H*(x-T_a) + k_E*(RH_s*0.61078*exp(17.27*x/(x+237.3)) - RH*0.61078*exp(17.27*T_a/(T_a+237.3))))

}

#solve for x using uhi dataframe
solve <- function(x) pmap_dbl(uhi, fun)
uniroot.all(solve(), c(0,100))
 

```


*PART 2* 

Use the surface energy balance function with our Austin met datato calculate cooling degree days and estimated reduction in residential energy temp for our hourly met data 

*using mean, max, and minimum esimates from the data until we get the function to read the entire met file*


A. Start by getting daily high temp

```{r}
#use hourly dataset to calculate daily high temp
daily_high <- hourly %>% 
  group_by(id) %>%
  filter(Tair == max(Tair)) %>% 
  distinct(id, .keep_all = TRUE)
```

B. Use daily high temps to estimate max, min and mean temps for the summer

```{r}

#find maximum, minimum, and average temperature in hourly met data from 2017 

max_temp <- max(daily_high$Tair, na.rm = TRUE)

#maximum temperature was 36.111 degrees C

min_temp <- min(daily_high$Tair, na.rm = TRUE)

#minimum temp was 10.555 degrees C 

mean_temp <- mean(daily_high$Tair, na.rm = TRUE)

#mean temp was 24.19 degrees C  

#plug these numbers back into the forumal in part 1, with 100% irrigation, 50% irrigation, and 0% irrigation
```

C. Plug these numbers into the surface energy balance function (for max, mean, and min estimates)

Calculating maximum temperature estimates (changing k_E and RH_s) 
*I changed k_E and RS together, assuming that a completely dry surafce would not have a relative surface humidity of 1, and vice versa* 
```{r}

#Plug our maximum temp (max_temp = 41.11) number back into the formula in part 1, with 100% k_E (irrigation) and RH_s (relative surface humidity) of 1

fun <- function(x, T_a = max_temp, k_E = 100, RH_s = 1){ (-Q_av + eps*sigma*(x+273.15)^4 + k_H*(x-T_a) + k_E*(RH_s*0.61078*exp(17.27*x/(x+237.3)) - RH*0.61078*exp(17.27*T_a/(T_a+237.3))))

}

range = c(0,100)
max_cool = uniroot.all(fun, range)
max_cool 

#Calculate with 50% k_E and RH_s of 0.5

fun <- function(x, T_a = max_temp, k_E = 50, RH_s = 1){ (-Q_av + eps*sigma*(x+273.15)^4 + k_H*(x-T_a) + k_E*(RH_s*0.61078*exp(17.27*x/(x+237.3)) - RH*0.61078*exp(17.27*T_a/(T_a+237.3))))

}

range = c(0,100)
max_irrigation_50 = uniroot.all(fun, range)
max_irrigation_50 


#Calculate with 0% k_E and RH_s of 0


fun <- function(x, T_a = max_temp, k_E = 0, RH_s = 1){ (-Q_av + eps*sigma*(x+273.15)^4 + k_H*(x-T_a) + k_E*(RH_s*0.61078*exp(17.27*x/(x+237.3)) - RH*0.61078*exp(17.27*T_a/(T_a+237.3))))

}

range = c(0,100)
max_irrigation_0 = uniroot.all(fun, range)
max_irrigation_0 



```

Calculating mean temperature estimates 
*I changed k_E and RS together, assuming that a completely dry surafce would not have a relative surface humidity of 1, and vice versa* 

```{r}
#Plug our maximum temp (mean_temp = 33.84) number back into the formula in part 1, with 100% k_E (irrigation) and RH_s (relative surface humidity) of 1

fun <- function(x, T_a = mean_temp, k_E = 100, RH_s = 1){ (-Q_av + eps*sigma*(x+273.15)^4 + k_H*(x-T_a) + k_E*(RH_s*0.61078*exp(17.27*x/(x+237.3)) - RH*0.61078*exp(17.27*T_a/(T_a+237.3))))

}

range = c(0,100)
mean_cool = uniroot.all(fun, range)
mean_cool 

#Calculate with 50% k_E and RH_s of 0.5

fun <- function(x, T_a = mean_temp, k_E = 50, RH_s = .5){ (-Q_av + eps*sigma*(x+273.15)^4 + k_H*(x-T_a) + k_E*(RH_s*0.61078*exp(17.27*x/(x+237.3)) - RH*0.61078*exp(17.27*T_a/(T_a+237.3))))

}

range = c(0,100)
mean_irrigation_50 = uniroot.all(fun, range)
mean_irrigation_50 



#Calculate with 0% k_E and RH_s of 0


fun <- function(x, T_a = mean_temp, k_E = 0, RH_s = 0){ (-Q_av + eps*sigma*(x+273.15)^4 + k_H*(x-T_a) + k_E*(RH_s*0.61078*exp(17.27*x/(x+237.3)) - RH*0.61078*exp(17.27*T_a/(T_a+237.3))))

}

range = c(0,100)
mean_irrigation_0 = uniroot.all(fun, range)
mean_irrigation_0 
```


Calculating minimum temperature estimates

```{r}
#Plug our maximum temp (min_temp = 22.78) number back into the formula in part 1, with 100% k_E (irrigation) and RH_s (relative surface humidity) of 1

fun <- function(x, T_a = min_temp, k_E = 100, RH_s = 1){ (-Q_av + eps*sigma*(x+273.15)^4 + k_H*(x-T_a) + k_E*(RH_s*0.61078*exp(17.27*x/(x+237.3)) - RH*0.61078*exp(17.27*T_a/(T_a+237.3))))

}

range = c(0,100)
min_cool = uniroot.all(fun, range)
min_cool 

#Calculate with 50% k_E and RH_s of 0.5

fun <- function(x, T_a = min_temp, k_E = 50, RH_s = .5){ (-Q_av + eps*sigma*(x+273.15)^4 + k_H*(x-T_a) + k_E*(RH_s*0.61078*exp(17.27*x/(x+237.3)) - RH*0.61078*exp(17.27*T_a/(T_a+237.3))))

}

range = c(0,100)
min_irrigation_50 = uniroot.all(fun, range)
min_irrigation_50 


#Calculate with 0% k_E and RH_s of 0


fun <- function(x, T_a = min_temp, k_E = 0, RH_s = 0){ (-Q_av + eps*sigma*(x+273.15)^4 + k_H*(x-T_a) + k_E*(RH_s*0.61078*exp(17.27*x/(x+237.3)) - RH*0.61078*exp(17.27*T_a/(T_a+237.3))))

}

range = c(0,100)
min_irrigation_0 = uniroot.all(fun, range)
min_irrigation_0 
```

D. Cacluate cooling degree days 

```{r}
#use daily high cooling baseline using 75.3 as found in NREL study and cooling coefficient as found in Deschenes et al (SUBJEcT TO CHANGE)

#convert baseline to celsius 

cooling_baseline = ((75.3 -32)*5)/9

cooling_coefficient = 0.00139

#calculate current number of cooling days 

mean_current_cooling = (mean_temp - cooling_baseline)*152

max_current_cooling = (max_temp - cooling_baseline)*152

min_current_cooling = (min_temp - cooling_baseline)*152

#calculate new cooling days based on mean temp

mean_new_cooling = (mean_cool - cooling_baseline)*152 

#calculate new cooling days based on high temp

max_new_cooling = (max_cool - cooling_baseline)*152 

#calculate new cooling days based on low temp

min_new_cooling = (min_cool - cooling_baseline)*152

#difference in mean number of cooling days from current

difference_mean_days = mean_current_cooling-mean_new_cooling

mean_cooling_estimate = difference_mean_days/100 *cooling_coefficient


#difference in max number of cooling days from current

difference_max_days = max_current_cooling - max_new_cooling

max_cooling_estimate = difference_max_days/100 *cooling_coefficient

#difference in min number of cooling days from current

difference_min_days = min_current_cooling - min_new_cooling

min_cooling_estimate = difference_min_days/100 * cooling_coefficient 
```

E. Apply to Austin's annual energy consumption (will also look @ monthly consumption)
```{r}

#get city of Austin residential energy data from open data portal 

#Average annual kwH for 2017 

annual =10156 

#residential energy rates for 2017 

rate = 0.1051

mean_energy_reduction = annual*mean_cooling_estimate

max_energy_reduction = annual*max_cooling_estimate

#aggregate these estimates across 400 homes 

ag_mean_energy_reduction = mean_energy_reduction*400

ag_max_energy_reduction = max_energy_reduction*400 

#use $/kwH to estimate savings at the home and aggregate level 

ind_mean_savings = mean_energy_reduction * rate

ind_max_savings = max_energy_reduction * rate 

ag_mean_savings = ag_mean_energy_reduction *rate 

ag_max_savings = ag_max_energy_reduction * rate


```

*PART 3* 

Calculating savings monthly instead of annually 

Repeat the steps from part 2, but for each month instead of annually, to get estimates of reduced energy savings each month instead of an annual figure 
```{r}
#get mean high temps from each month 

may_temp <- daily_high %>% 
  filter(id < 152)

may_mean_temp <- mean(may_temp$Tair, na.rm = TRUE)

june_temp <- daily_high %>% 
  filter(151 < id, id < 182)

june_mean_temp <- mean(june_temp$Tair, na.rm = TRUE)

july_temp <- daily_high %>% 
  filter(181 < id, id < 213)

july_mean_temp <- mean(july_temp$Tair, na.rm = TRUE)

august_temp <- daily_high %>% 
  filter(212 < id, id < 243)

august_mean_temp <- mean(august_temp$Tair, na.rm = TRUE)

sept_temp <- daily_high %>% 
  filter(242 < id)

sept_mean_temp <- mean(sept_temp$Tair, na.rm = TRUE)

```

Plug these numbers into the surface energy balance equation 
```{r}
fun <- function(x, T_a = sept_mean_temp, k_E = 100, RH_s = 1){ (-Q_av + eps*sigma*(x+273.15)^4 + k_H*(x-T_a) + k_E*(RH_s*0.61078*exp(17.27*x/(x+237.3)) - RH*0.61078*exp(17.27*T_a/(T_a+237.3))))

}

range = c(0,100)
monthly_temp = uniroot.all(fun, range)
monthly_temp


#name each monthly average surface temp

may_mean_cool = 29.68596 
june_mean_cool = 32.29015 
july_mean_cool = 35.30106 
august_mean_cool = 32.81764 
sept_mean_cool = 30.48862 


```

Visualize the results 
```{r}
months <- c("May", "June", "July", "August", "September")

Actual<- c(may_mean_temp, june_mean_temp, july_mean_temp, august_mean_temp, sept_mean_temp)

Predicted <- c(may_mean_cool, june_mean_cool, july_mean_cool, august_mean_cool, sept_mean_cool)

temp_df <- data.frame(months, temperature_actual, temperature_predicted)



melted_df <- melt(temp_df, id = "months") 

temp_graph <- ggplot(melted_df, aes(x = months, y = value)) +
                       geom_col(aes(fill = variable), position = "dodge") +
                      scale_fill_manual(values = c("#33CCCC", "#00AFBB")) +
                      theme_classic() +
  labs(x = "Month", y = "Temperature (Degrees C)", title = "Actual vs. Predicted Temperature (Summer 2017)") 

temp_graph
```

Calculate cooling days for each month 
```{r}

#cooling days for may 

cooling_days_may_current = (may_mean_temp - cooling_baseline)*31

cooling_days_may_predicted = (may_mean_cool - cooling_baseline)*31

difference_may = (cooling_days_may_current - cooling_days_may_predicted)

#cooling days for June 

cooling_days_june_current = (june_mean_temp - cooling_baseline)*30

cooling_days_june_predicted = (june_mean_cool - cooling_baseline)*30

difference_june = (cooling_days_june_current - cooling_days_june_predicted)

#cooling days for July 

cooling_days_july_current = (july_mean_temp - cooling_baseline)*31

cooling_days_july_predicted = (july_mean_cool - cooling_baseline)*31

difference_july = (cooling_days_july_current - cooling_days_july_predicted)

#cooling days for August 

cooling_days_august_current = (august_mean_temp - cooling_baseline)*31

cooling_days_august_predicted = (august_mean_cool - cooling_baseline)*31

difference_august = (cooling_days_august_current - cooling_days_august_predicted)

#cooling days for September 

cooling_days_sept_current = (sept_mean_temp - cooling_baseline)*30

cooling_days_sept_predicted = (sept_mean_cool - cooling_baseline)*30

difference_sept = (cooling_days_sept_current - cooling_days_sept_predicted)

```

Calculate estimated 

```{r}
#get monthly cooling coefficient instead of annual 

monthly_cooling_coefficient = cooling_coefficient*12

#calculate new cooling degree days and impact on energy 

#may

may_cooling_estimate = difference_may/100 *monthly_cooling_coefficient

#june 

june_cooling_estimate = difference_june/100 *monthly_cooling_coefficient

#july 

july_cooling_estimate = difference_july/100 *monthly_cooling_coefficient

#august 

august_cooling_estimate = difference_august/100 *monthly_cooling_coefficient

#september 

sept_cooling_estimate = difference_sept/100 *monthly_cooling_coefficient
```

Apply this to Austin average monthly energy consumption 
```{r}
#get city of Austin residential energy data from open data portal 

#Average kwH for May 2017

may = 708 

#Average kwH for June 2017

june = 959

#Average kwH for July 2017

july = 1246 

#Average kwH for August 2017

august = 1277 

#average kwH for September 2017 

sept = 1128 


#caclulate individual and aggregate energy reduction for each month

#May 
may_energy_reduction = may*may_cooling_estimate

ag_may = may_energy_reduction*400

#June

june_energy_reduction = june*june_cooling_estimate

ag_june = june_energy_reduction*400 

#July 

july_energy_reduction = july*july_cooling_estimate 

ag_july = july_energy_reduction*400 

#August 

august_energy_reduction = august*august_cooling_estimate

ag_august = august_energy_reduction*400 

#Sept 

sept_energy_reduction = sept*sept_cooling_estimate

ag_sept = sept_energy_reduction*400 


#add them up to check work from part 2 
total_energy_reduction = ag_may + ag_june + ag_july + ag_august + ag_sept 
```

Visualize this 
```{r}
Energy <- c(may, june, july, august, sept)

Reduction <- c(may_energy_reduction, june_energy_reduction, july_energy_reduction, august_energy_reduction, sept_energy_reduction)

energy_df <- data.frame(months, Reduction, Energy)

melted_energy <- melt(energy_df, id = "months") 

temp_graph <- ggplot(melted_energy, aes(x = months, y = value)) +
                       geom_col(aes(fill = variable)) +
                      scale_fill_manual(values = c("darkred", "darkgreen")) +
                      theme_classic() +
  labs(x = "Month", y = "Energy Consumption (kWh", title = "Potential Reduction in Energy Use") 

temp_graph

```

Caclulate monthly energy savings using rate from part 2
```{r}
#calculate individual and aggregate savings for each month 

#May 

may_ind_savings = may_energy_reduction * rate

may_ag_savings = ag_may * rate 

#June 

june_ind_savings = june_energy_reduction *rate 

june_ag_savings = ag_june* rate 

#July 

july_ind_savings = july_energy_reduction * rate 

july_ag_savings = ag_july * rate 

#August 

august_ind_savings = august_energy_reduction * rate 

august_ag_savings = ag_august * rate 

#September 

sept_ind_savings = sept_energy_reduction * rate 

sept_ag_savings = ag_sept * rate 


#add them up to check work from part 2 

total_savings = may_ag_savings + june_ag_savings + july_ag_savings + august_ag_savings + sept_ag_savings

#monthly estimates are a bit higher than annual estimates 
```

Visualize energy savings 
```{r}

```

