---
title: "GP_UHI"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages 
```{r}
library(rootSolve)
library(tidyverse)
```


load csv 
```{r}
uhi <- read_csv("uhi_test.csv")

```

### Surface Energy Balance - Solving for surface temperature 

1. Run this code first to set the variables. 

Note: For setting the variables, we've set the initial values as the same for the parking lot example in your assignment. 

```{r vars}
T_a = 33 # ambient temp in Celsius 
alpha = 0.2 # albedo 
eps = 0.95 # emissivity 
sigma = 5.6704*10^-8 #W/m^2/K^4 stefan-boltzmann constant
k_H = 21 # sensible heat conductivity [J/deg-C]
k_E = 0 # 0 for non-wetted surface, 100 for fully wet. [J/kPa] Latent energy only present if surface is wet

RH = 0.7 # Relative humidity of atmosphere, [0-1]
RH_s = 0 # Relative humidity of the surface (assume 1 is wet/irrigated, 0 is dry)

K_inc = 750 # W/m^2, incoming shortwave radiation
#L_inc = eps * sigma * T_a^4  # W/m^2, incoming longwave radiation
Q_av = 0.9 * (K_inc*(1-alpha) + eps * sigma * T_a^4) # W/m^2 

```

2. Run this to solve for the surface temperature 

```{r function}
# temperature is represented as x 
fun <- function(x){ (-Q_av + eps*sigma*(x+273.15)^4 + k_H*(x-T_a) + k_E*(RH_s*0.61078*exp(17.27*x/(x+237.3)) - RH*0.61078*exp(17.27*T_a/(T_a+237.3))))

}

# solve function for temp, within the set range 
range = c(0,100)
temp_s = uniroot.all(fun, range)
temp_s


```
T_a = 33 # ambient temp in Celsius 
alpha = 0.2 # albedo 
eps = 0.95 # emissivity 
sigma = 5.6704*10^-8 #W/m^2/K^4 stefan-boltzmann constant
k_H = 21 # sensible heat conductivity [J/deg-C]
k_E = 0 # 0 for non-wetted surface, 100 for fully wet. [J/kPa] Latent energy only present if surface is wet

RH = 0.7 # Relative humidity of atmosphere, [0-1]
RH_s = 0 # Relative humidity of the surface (assume 1 is wet/irrigated, 0 is dry)

K_inc = 750 # W/m^2, incoming shortwave radiation
L_inc = eps * sigma * T_a^4  # W/m^2, incoming longwave radiation
Q_av = 0.9 * (K_inc*(1-alpha) + L_inc) # W/m^2 



```{r function}
range = c(0,100)
# temperature is represented as x 
fun = function(T_a){ -(0.9 * (K_inc*(1-alpha) + eps * sigma * (T_a)^4) + eps*sigma*(x+273.15)^4 + k_H*(x-(T_a)) + k_E*(RH_s*0.61078*exp(17.27*x/(1+237.3)) - RH*0.61078*exp(17.27*(T_a)/(T_a+237.3))))

return(x)
}


uhi$temp = fun(T_a = uhi$Tair)

# solve function for temp, within the set range 
range = c(0,100)
temp_s = uniroot.all(fun, range)
temp_s
```

#' Power Required by Speed
#'
#' This function determines the power required to keep a vehicle moving at 
#' a given speed
#' @param cdrag coefficient due to drag default=0.3 
#' @param crolling coefficient due to rolling/friction default=0.015
#' @param v vehicle speed (m/2)
#' @param m vehicle mass (kg)
#' @param A area of front of vehicle (m2)
#' @param g acceleration due to gravity (m/s) default=9.8
#' @param pair (kg/m3) default =1.2
#' @return power (W)

autopower = function(V, m, A, cdrag=0.3, crolling=0.015,pair=1.2,g=9.8) {
  P = crolling*m*g*V + 1/2*A*pair*cdrag*V**3
  return(P)
}


possible_cars$power = autopower(V=speed_base, A = possible_cars$area, m=possible_cars$area)


With the given inputs from step 1, the surface temperature is `r format(round(temp_s, 2), nsmall = 2)` degrees C. 
