---
title: "GP_UHI"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

***ISSUES*** 
 The function is not solved for x. With fixed values, I can use uniroot to find the solution, but this doesnt't work when I try to use values from a dataframe. 
 
Line 72 is my latest attempt at using our dataframe


Load packages 
```{r}
library(rootSolve)
library(tidyverse)
library(purrr)
library(dplyr)

```


load csv 
```{r}
uhi <- read_csv("uhi_test.csv")
hourly <- read_csv("hourly_2017_met.csv")

```

### Surface Energy Balance - Solving for surface temperature 

1. Run this code first to set the variables. These variables were given in the problem


```{r vars}
T_a = 10.555# ambient temp in Celsius 
alpha = 0.2 # albedo 
eps = 0.95 # emissivity 
sigma = 5.6704*10^-8 #W/m^2/K^4 stefan-boltzmann constant
k_H = 21 # sensible heat conductivity [J/deg-C]
k_E = 100 # 0 for non-wetted surface, 100 for fully wet. [J/kPa] Latent energy only present if surface is wet

RH = 0.7 # Relative humidity of atmosphere, [0-1]
RH_s = 1 # Relative humidity of the surface (assume 1 is wet/irrigated, 0 is dry)

K_inc = 750 # W/m^2, incoming shortwave radiation
#L_inc = eps * sigma * T_a^4  # W/m^2, incoming longwave radiation
Q_av = 0.9 * (K_inc*(1-alpha) + eps * sigma * T_a^4) # W/m^2 

```

2. Run this to solve for the surface temperature (with fixed values)

```{r function}
# temperature is represented as x 
fun <- function(x){ (-Q_av + eps*sigma*(x+273.15)^4 + k_H*(x-T_a) + k_E*(RH_s*0.61078*exp(17.27*x/(x+237.3)) - RH*0.61078*exp(17.27*T_a/(T_a+237.3))))

}

# solve function for temp, within the set range 
range = c(0,100)
temp_s = uniroot.all(fun, range)
temp_s


```


3. Try inputting dataframe uhi to function 

```{r function}
# temperature is represented as x 
fun <- function(alpha, eps, sigma, RH, RH_s, T_a, K_inc, k_H, k_E, Q_av){ (-Q_av + eps*sigma*(x+273.15)^4 + k_H*(x-T_a) + k_E*(RH_s*0.61078*exp(17.27*x/(x+237.3)) - RH*0.61078*exp(17.27*T_a/(T_a+237.3))))

}

#solve for x using uhi dataframe
solve <- function(x) pmap_dbl(uhi, fun)
uniroot.all(solve(), c(0,100))
 

```

Use the surface energy balance function to calculate cooling degree days and estimated reduction in residential energy temp for our hourly met data 
*using fixed esimates until we get the function to read the met file*


Start by getting daily high temp

```{r}
#use hourly dataset to calculate daily high temp
daily_high <- hourly %>% 
  group_by(id) %>%
  filter(Tair == max(Tair)) %>% 
  distinct(id, .keep_all = TRUE)
```

Use daily high temps to estimate max, min and mean temps for the summer

```{r}

#find maximum, minimum, and average temperature in hourly met data from 2017 

max_temp <- max(daily_high$Tair, na.rm = TRUE)

#maximum temperature was 36.111 degrees C

min_temp <- min(daily_high$Tair, na.rm = TRUE)

#minimum temp was 10.555 degrees C 

mean_temp <- mean(daily_high$Tair, na.rm = TRUE)

#mean temp was 24.19 degrees C  

#plug these numbers back into the forumal in part 1, with 100% irrigation, 50% irrigation, and 0% irrigation
```

Plug these numbers into the surface energy balance function (for max, mean, and min estimates)

Calculating maximum temperature estimates (changing k_E and RH_s) 
*I changed k_E and RS together, assuming that a completely dry surafce would not have a relative surface humidity of 1, and vice versa* 
```{r}

#Plug our maximum temp (max_temp = 41.11) number back into the formula in part 1, with 100% k_E (irrigation) and RH_s (relative surface humidity) of 1

fun <- function(x, T_a = max_temp, k_E = 100, RH_s = 1){ (-Q_av + eps*sigma*(x+273.15)^4 + k_H*(x-T_a) + k_E*(RH_s*0.61078*exp(17.27*x/(x+237.3)) - RH*0.61078*exp(17.27*T_a/(T_a+237.3))))

}

range = c(0,100)
max_cool = uniroot.all(fun, range)
max_cool 

#Calculate with 50% k_E and RH_s of 0.5

fun <- function(x, T_a = max_temp, k_E = 50, RH_s = 1){ (-Q_av + eps*sigma*(x+273.15)^4 + k_H*(x-T_a) + k_E*(RH_s*0.61078*exp(17.27*x/(x+237.3)) - RH*0.61078*exp(17.27*T_a/(T_a+237.3))))

}

range = c(0,100)
max_irrigation_50 = uniroot.all(fun, range)
max_irrigation_50 


#Calculate with 0% k_E and RH_s of 0


fun <- function(x, T_a = max_temp, k_E = 0, RH_s = 1){ (-Q_av + eps*sigma*(x+273.15)^4 + k_H*(x-T_a) + k_E*(RH_s*0.61078*exp(17.27*x/(x+237.3)) - RH*0.61078*exp(17.27*T_a/(T_a+237.3))))

}

range = c(0,100)
max_irrigation_0 = uniroot.all(fun, range)
max_irrigation_0 



```

Calculating average mean estimates 
*I changed k_E and RS together, assuming that a completely dry surafce would not have a relative surface humidity of 1, and vice versa* 

```{r}
#Plug our maximum temp (mean_temp = 33.84) number back into the formula in part 1, with 100% k_E (irrigation) and RH_s (relative surface humidity) of 1

fun <- function(x, T_a = mean_temp, k_E = 100, RH_s = 1){ (-Q_av + eps*sigma*(x+273.15)^4 + k_H*(x-T_a) + k_E*(RH_s*0.61078*exp(17.27*x/(x+237.3)) - RH*0.61078*exp(17.27*T_a/(T_a+237.3))))

}

range = c(0,100)
mean_cool = uniroot.all(fun, range)
mean_cool 

#Calculate with 50% k_E and RH_s of 0.5

fun <- function(x, T_a = mean_temp, k_E = 50, RH_s = .5){ (-Q_av + eps*sigma*(x+273.15)^4 + k_H*(x-T_a) + k_E*(RH_s*0.61078*exp(17.27*x/(x+237.3)) - RH*0.61078*exp(17.27*T_a/(T_a+237.3))))

}

range = c(0,100)
mean_irrigation_50 = uniroot.all(fun, range)
mean_irrigation_50 



#Calculate with 0% k_E and RH_s of 0


fun <- function(x, T_a = mean_temp, k_E = 0, RH_s = 0){ (-Q_av + eps*sigma*(x+273.15)^4 + k_H*(x-T_a) + k_E*(RH_s*0.61078*exp(17.27*x/(x+237.3)) - RH*0.61078*exp(17.27*T_a/(T_a+237.3))))

}

range = c(0,100)
mean_irrigation_0 = uniroot.all(fun, range)
mean_irrigation_0 
```


Calculating minimum temperature estimates

```{r}
#Plug our maximum temp (min_temp = 22.78) number back into the formula in part 1, with 100% k_E (irrigation) and RH_s (relative surface humidity) of 1

fun <- function(x, T_a = min_temp, k_E = 100, RH_s = 1){ (-Q_av + eps*sigma*(x+273.15)^4 + k_H*(x-T_a) + k_E*(RH_s*0.61078*exp(17.27*x/(x+237.3)) - RH*0.61078*exp(17.27*T_a/(T_a+237.3))))

}

range = c(0,100)
min_cool = uniroot.all(fun, range)
min_cool 

#Calculate with 50% k_E and RH_s of 0.5

fun <- function(x, T_a = min_temp, k_E = 50, RH_s = .5){ (-Q_av + eps*sigma*(x+273.15)^4 + k_H*(x-T_a) + k_E*(RH_s*0.61078*exp(17.27*x/(x+237.3)) - RH*0.61078*exp(17.27*T_a/(T_a+237.3))))

}

range = c(0,100)
min_irrigation_50 = uniroot.all(fun, range)
min_irrigation_50 


#Calculate with 0% k_E and RH_s of 0


fun <- function(x, T_a = min_temp, k_E = 0, RH_s = 0){ (-Q_av + eps*sigma*(x+273.15)^4 + k_H*(x-T_a) + k_E*(RH_s*0.61078*exp(17.27*x/(x+237.3)) - RH*0.61078*exp(17.27*T_a/(T_a+237.3))))

}

range = c(0,100)
min_irrigation_0 = uniroot.all(fun, range)
min_irrigation_0 
```

Cacluate cooling degree days 

```{r}
#use daily high cooling baseline using 75.3 as found in NREL study and cooling coefficient as found in Deschenes et al (SUBJEcT TO CHANGE)

#convert baseline to celsius 

cooling_baseline = ((75.3 -32)*5)/9

cooling_coefficient = 0.0139

#calculate current number of cooling days 

mean_current_cooling = (mean_temp - cooling_baseline)*152

max_current_cooling = (max_temp - cooling_baseline)*152

min_current_cooling = (min_temp - cooling_baseline)*152

#calculate new cooling days based on mean temp

mean_new_cooling = (mean_cool - cooling_baseline)*152 

#calculate new cooling days based on high temp

max_new_cooling = (max_cool - cooling_baseline)*152 

#calculate new cooling days based on low temp

min_new_cooling = (min_cool - cooling_baseline)*152

#difference in mean number of cooling days from current

difference_mean_days = mean_current_cooling-mean_new_cooling

mean_cooling_estimate = difference_mean_days/100 *cooling_coefficient


#difference in max number of cooling days from current

difference_max_days = max_current_cooling - max_new_cooling

max_cooling_estimate = difference_max_days/100 *cooling_coefficient

#difference in min number of cooling days from current

difference_min_days = min_current_cooling - min_new_cooling

min_cooling_estimate = difference_min_days/100 * cooling_coefficient 
```

Apply to Austin's annual energy consumption (will also look @ monthly consumption)
```{r}

#get city of Austin residential energy data from open data portal 

#Average annual kwH for 2017 

annual =10156 

#residential energy rates for 2017 

rate = 0.1051

mean_energy_reduction = annual*mean_cooling_estimate

max_energy_reduction = annual*max_cooling_estimate

#aggregate these estimates across 400 homes 

ag_mean_energy_reduction = mean_energy_reduction*400

ag_max_energy_reduction = max_energy_reduction*400 

#use $/kwH to estimate savings at the home and aggregate level 

ind_mean_savings = mean_energy_reduction * rate

ind_max_savings = max_energy_reduction * rate 

ag_mean_savings = ag_mean_energy_reduction *rate 

ag_max_savings = ag_max_energy_reduction * rate


#calculate a monthly savings instead?? could be usefuk if we don't use entire summertime distribution 

#Average kwH for May 2017

may = 708 

#Average kwH for June 2017

june = 959

#Average kwH for July 2017

july = 1246 

#Average kwH for August 2017

august = 1277 

#average kwH for September 2017 

september = 1128 


#calculate estimated savings 


```


